Model-order reduction of non-linear time dependent neutronics via Galerkin projection and DEIM

### 2022-02-24 paper/interactnlmsample.tex
updated text

\usepackage{algorithm,algorithmic}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{booktabs,tabularx}
\DeclareMathOperator{\argmaxH}{argmax} 
\DeclareMathOperator*{\argmin}{argmin}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand\abs[1]{\left|#1\right|}

@@

Modeling and simulation of nuclear reactors is crucial to predict their behavior under different conditions and scenarios and hence ensure reliable design and safety limits.
The mathematical models are described by a set of equations that is challenging to solve analytically due to the geometric complexity of nuclear reactors except for simplified cases.
Alternatively, these systems are solved by employing numerical discretization that rely on discretizing the system along each independent variable and the solution is provided only on these discrete points.
To improve the method accuracy, the system needs to be solved over a very fine mesh.
Although this results in a more reliable solution, it leads to a growing space dimension and accordingly large computational cost.
This problem of the computational cost becomes inevitably prohibitive with applications that require repetitive execution of the simulation such as uncertainty quantification, and design optimization.
In such studies, the high-fidelity model has to be solved at multiple parameter points which could be very large.
Hence, surrogate models that are cheap, yet accurate, are sought to provide a rapid and reliable approximation of the solution of these systems.
The premise is that this lower-order model will capture the essential features of the system  and preserve the input-output mapping as much as possible.

@@

\section{Methods Overview}
\subsection{POD-Galerkin Projection}\label{class}

Consider the following dynamic system that results from discretizing a partial differential equation using a numerical scheme such as the finite difference method:
\begin{equation}
\frac{d\mathbf{y}(t)}{dt} = \mathbf{A}\mathbf{y}(t) + \mathbf{f}(\mathbf{y}(t)),  \quad\quad \mathbf{y(0) = \mathbf{y_0}} \, ,
\label{eq:FOM}
\end{equation}
where $\mathbf{y} \in \mathbb{R}^{n}$ is the state variable and $\mathbf{A} \in \mathbb{R}^{n\times n}$ is the operator resulting from the system discretization.
In general, this method involves two main steps. 
First, the state variable $\mathbf{y}$ is expanded in terms of a trial subspace basis, which is the POD space in this case, where each basis vector is weighted by a temporal coefficient
\begin{equation}
\mathbf{y} \approx \mathbf{\Phi} \hat{\mathbf{y}}(t) \, ,
\label{eq:galerkin expansion}
\end{equation}
where $\mathbf{\hat{y}(t)} \in \mathbb{R}^{r}$ is a vector of the temporal coefficients.
Second, the system is projected onto a test subspace, which is the same as the trial subspace in the Galerkin projection procedures.
By multiplying the system on the left by $\mathbf{\Phi}^T$ and making use of the orthogonality of the POD basis (i.e, $\mathbf{\Phi}^T\mathbf{\Phi} = \mathbf{I}$), we end up with 
\begin{equation}
\frac{d\mathbf{\hat{y}}}{dt} = \mathbf{A}_r\mathbf{\hat{y}} + \mathbf{\Phi}^T\mathbf{f}(\mathbf{\Phi}\mathbf{\hat{y}}) \qquad \mathbf{\hat{y}}(0) = \mathbf{\Phi}^T\mathbf{y}_0 \, ,
\label{eq:ROM}
\end{equation}
where $\mathbf{A}_r \in \mathbb{R}^{r\times r} = \mathbf{\Phi}^T\mathbf{A}\mathbf{\Phi}.$
Thus, we are left with  $r$ ODE's instead of $n$ ODE's, which makes the resulting ROM more computationally efficien, and, once it is solved for $\mathbf{\hat{y}}$, the full order solution can be computed by Eq.~\ref{eq:galerkin expansion}.
Note that the POD basis is chosen here for its optimality property.
However, different methods have been used in the literature, such as the eigenmodes analysis and the reduced-basis method \cite{rozza2008reduced, hesthaven2016certified, veroy2003posteriori}.
Moreover, the test basis and the trial basis could be different, in which case the method is called Petrov-Galerkin projection.
In such cases, unless the test and the trial basis are bi-orthogonal, there will be an additional computational cost associated with inverting the matrix resulting from multiplying the left hand side of Eq.~\ref{eq:FOM} by the test basis.

\subsection{Discrete Impricical Interpolation Method \emph{\LaTeX}}

\label{sec:DEIM}
Recently, the DEIM \cite{chaturantabut2010nonlinear} was proposed to approximate the non-linearity in model order reduction.
The method is a discrete variant of the Empirical Interpolation Method (EIM) \cite{barrault2004empirical} that was originally proposed to approximate
non-affine coefficient functions using a reduced-basis expansion to facilitate the online/offline decomposition in finite element methods.
The DEIM is proposed for the same goal for different discretizations of the full-order model.
The method relies on two main ingredients: (1) a reduced space that approximates the nonlinear function, and (2) a set of spatial interpolation indices.
To illustrate the method, let $\tau$ denote time or a model parameter.
The nonlinear function $\mathbf{f}(\textbf{y}(\tau))$ is approximated by projecting it onto a subspace that approximates the function space
\begin{equation}
\mathbf{f}(\textbf{y}(\tau)) = \mathbf{\Psi}\boldsymbol{\theta}(\tau)\, ,
\end{equation}
where $\mathbf{\Psi} \in \mathbb{R}^{n\times k}$ is a low-dimensional subspace with $k \ll n$, and $\boldsymbol{\theta}(\tau) \in \mathbb{R}^k$ is a vector of the corresponding coefficients.
POD is used to find the subspace $\mathbf{\Psi}$ using snapshots of the function (i.e, $[\textbf{y}(\tau_1), ....., \textbf{y}(\tau_m)]$).
Note that these snapshots are computed at the same points used for the state variable snapshots required for the reduced model; hence this step will not introduce any additional cost other than computing the SVD.
In order to compute $\theta(\tau)$, a few spatial indices are selected iteratively using a greedy search following Algorithm \ref{alg:DEIM}.
The basic idea is to select $k$ rows of the basis matrix $\mathbf{\Psi}$.
The algorithm takes as an input the basis vectors of the subspace $\mathbf{\Psi}$.
The first index is selected where the entry of the first input basis has the largest magnitude.
Each other index $\mathcal{I}_i$ is selected where the approximation of the corresponding input basis vector $\mathbf{\Psi_i}$ in the subspace constructed by interpolating the basis $[\mathbf{\Psi}_1, ....\mathbf{\Psi}_{i-1}]$ at the previously selected indices is worst.
Once the indices are selected, the matrix $\mathbf{\Psi}_k \in \mathbb{R}^{k\times k}$ is formed such that
\begin{equation}
\mathbf{\Psi}_k = \mathbf{P}^T\mathbf{\Psi}\, ,
\end{equation}
where $\mathbf{P} = [e_{z1}, e_{z2}, ....,e_{zk}] \in \mathbb{R}^{n\times k}$ and $e_{zi} = [0,.., 0, \underbrace{1}_{{zi}}, 0..., 0]^\text{T} \in \mathbb{R}^{n}$
is the  ${zi}$th column of the identity matrix.
The coefficients can be computed by solving
\begin{equation}
\mathbf{P}^T\mathbf{f(\tau)}=\mathbf{\Psi}_k \boldsymbol{\theta}(\tau)\, ,
\label{eq:DEIM coefficients}
\end{equation}
and the final function approximation is given by
\begin{equation}
\hat{\mathbf{f}}(\tau)\approx\mathbf{\Psi}( \mathbf{\Psi}_k)^{-1}\mathbf{P}^T\mathbf{f} (\tau)\, .
\end{equation}
Note that the matrix $\mathbf{P}$ does not need to be constructed explicitly, since the term $\mathbf{P}^T\mathbf{f}$ extracts only the elements of $\mathbf{f}$ that correspond to the interpolation indices.
Thus, at each time step, only $k$ elements $\mathbf{f}$ need to be evaluated.
The approximation error can be bounded as 
\begin{equation}
\|\mathbf{f} - \hat{\mathbf{f}} \| \le \|(\mathbf{P}\mathbf{\Psi})^{-1} \| \epsilon \, ,
\end{equation}
where
\begin{equation*}
\epsilon \approx \sigma_{k+1} \, .
\end{equation*}
Note that $\sigma_{k+1}$ is the $(k+1)th$ singular value coming from the SVD of the function snapshots matrix.
Full derivation of the error bound is given in Ref.~\cite{chaturantabut2010nonlinear}.

As explained in the previous section, the projection can be performed by computing the term $\mathbf{U}^T \mathbf{\Psi}$ once in an offline stage, while in the online stage, the vector $\boldsymbol{\theta}$ is computed by solving the linear system of dimension $k$ in Eq.~\ref{eq:DEIM coefficients}.
The key point is that the cost of the online computation depends on the reduced dimension $k$, not the original, high dimension $n$.
\noindent\begin{minipage}{\textwidth}
	\renewcommand\footnoterule{}
	\begin{algorithm}[H]
		\caption{DEIM}
		\label{alg:DEIM}
		\begin{algorithmic}[1]
			\REQUIRE Snapshots matrix: $\mathbf{X} = [\mathbf{f}_1, \mathbf{f}_2,  ...., \mathbf{f}_m] \in \mathbb{R}^{n\times m}$
			\ENSURE A reduced space $\mathbf{\Psi}$, set of indices $\mathcal{I}$
			\STATE $[\mathbf{\Psi}_1, \mathbf{\Psi}_2, ..., \mathbf{\Psi}_k ]$ = POD(${\mathbf{X}}, \epsilon$)\footnote{POD refers to Algorithm \ref{alg:pod}}
			\STATE $\mathcal{I}_1 = \argmaxH(\abs{\mathbf{\Psi}_1})$, $\mathcal{I}$ = $\{\mathcal{I}_1\}$
			\STATE $\mathbf{\Psi} = [\mathbf{\Psi}_1]$, $\mathbf{P} =[e_{I_1}]$
			\FOR {i=2 \TO i=k }
			\STATE Solve $(\mathbf{P}^T\mathbf{\Psi})\mathbf{c} = \mathbf{P}^T\mathbf{\Psi}_i$;
			\STATE $\mathcal{I}_i = \argmaxH{\abs{\mathbf{\Psi}_i - \mathbf{\Psi}\mathbf{c}}}$
			\STATE $\mathcal{I} = \mathcal{I} \cup \mathcal{I}_i$, $\mathbf{P} = \mathbf{P} \cup e_{I_i}$, ${\Psi = \Psi \cup \Psi_i}$
			\ENDFOR
		\end{algorithmic}
	\end{algorithm}
\end{minipage}

\subsection{Greedy POD}
A greedy sampling in the parameter space \cite{nguyen2009reduced} can be combined with POD to construct a global subspace that can be used in the framework of parametric ROMs.
The primary goal is to ensure that the parameter domain of interest (i.e, $\mathbb P$) is efficiently sampled and includes enough information such that it can be used or wide range of parameters including those of higher dimensions.
To construct this subspace, a training parameters dataset, $P_{train} \subset \mathbb P$, is prepared. 
The greedy subspace is initialized with the first $r_1$ POD basis computed from a randomly selected sample of the training data.
Next, this subspace is used to compute the ROM prediction error of the state variable for the rest of the training samples.
The sample at which the error is worst is selected, and the subspace is enriched by adding the first $r_1$ POD basis computed from the selected sample snapshots.
Selecting the samples with the worst errors ensures efficient sampling of the parameter space.
Intuitively, samples that are close to the previously chosen parameter points will have a smaller ROM error.
On the other hand, one expects that the ROM will fail to give an accurate prediction at samples that are far from the included points in the current subspace.
Thus, we keep adding new basis vectors until a maximum size of the subspace is reached or a user-specified error tolerance at a reference parameter point $\boldsymbol{\mu}_{\text{ref}}$ is met.
Note that each time a new basis vector is added, the subspace is orthonormalized.
Regarding the error criteria based on which a new sample is selected, there are two approaches. 
Ideally, the true error can be used leading to ``strong POD-greedy'' \cite{benner2017model}.
Alternatively, an error indicator can be used to alleviate the computational burden associated with evaluating the full-order-model. 
Here, we used the strong-POD greedy approach. 
The algorithm steps are shown in Algorithm \ref{alg:greedy pod}. 
The term $\Delta(\textbf{U},\boldsymbol{\mu}_\text{ref})$  is the error measure using the current reduced subspace at $\boldsymbol{\mu}_\text{ref}$, and $r$ is the maximum size of the greedy basis.

@@

\section{Benchmark Desription}
The LRA benchmark employs two energy groups and two delayed precursors and represents a 2-D quarter-core, BWR model subject to a control-rod ejection with adiabatic heating \cite{anl}.
Schematic layout of the problem is shown in Fig.~\ref{fig:lra core}, where region numbers indicate different materials and region R is the ejected rod.
%To produce time-dependent snapshots, the deterministic transport code Detran\footnote{https://github.com/robertsj/libdetran} was used to solve the time-dependent diffusion equations.
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=0.6]	
	\begin{scope}<->;

>>>>

\usepackage{algorithm,algorithmic}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{threeparttable}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{booktabs,tabularx}
\DeclareMathOperator{\argmaxH}{argmax} 
\DeclareMathOperator*{\argmin}{argmin}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand\abs[1]{\left|#1\right|}

@@

Modeling and simulation of reactor physics is crucial to understand the nuclear reactors behavior under different conditions and scenarios and hence ensure reliable design and well established safety limits.
The mathematical models are described by a set of equations that is challenging to solve analytically due to the geometric complexity of nuclear reactors except for simplified cases.
Alternatively, these systems are solved by employing numerical discretization that rely on discretizing the system along each independent variable and the solution is provided only on these discrete points.
To improve the method accuracy, the system needs to be solved over a very fine mesh.
Although this results in a more reliable solution, it leads to a growing space dimension and accordingly large computational cost.
This problem of the computational cost becomes inevitably prohibitive with applications that require repetitive execution of the simulation such as uncertainty quantification, and design optimization.
In such studies, the high-fidelity model has to be solved at multiple parameter points which could be very large.
Hence, surrogate models that are cheap, yet accurate, are sought to provide a rapid and reliable approximation of the solution of these systems.
The premise is that this lower-order model will capture the essential features of the system  and preserve the input-output mapping as much as possible.

@@

%--------------------------------------------------------------------------------------------------%
\section{Methods Overview}

\subsection{POD-Galerkin Projection}\label{class}

Consider the following dynamic system that results from discretizing a partial differential equation using a numerical scheme such as the finite difference method:
\begin{equation}
\frac{d\mathbf{y}(t)}{dt} = \mathbf{A}\mathbf{y}(t) + \mathbf{f}(\mathbf{y}(t)),  \quad\quad \mathbf{y(0) = \mathbf{y_0}} \, ,
\label{eq:FOM}
\end{equation}
where $\mathbf{y} \in \mathbb{R}^{n}$ is the state variable and $\mathbf{A} \in \mathbb{R}^{n\times n}$ is the operator resulting from the system discretization.
In general, this method involves two main steps.
First, a subspace is obtained using the POD which is a widely used reduction technique.
Commonly, the POD subspace is generated using the snapshots method \cite. 
A data matrix $\mathbf{Y}$ is constructed by collecting the system solutions at different time instances, then the singular value decomposition (SVD) is computed, i.e $\mathbf{Y} = \mathbf{U}\mathbf{\Sigma}\mathbf{V}$.
The POD subspace is the one spanned by the first $r$ vector of $\mathbf{U}$, where $r \ll n$.
There is no general \textit{a priori} rule based on which the rank is chosen; however one can set a threshold $\epsilon$  such that
\begin{equation}
\frac{\sum_{i=0}^r S_i}{\sum_{i=0}^m S_i} \ge \epsilon \, ,
\end{equation}
where S$i$ is the $i$th singular value.
For more details the reader might consult \cite{bibid}, 
Upon obtaining the reduced space, the state variable $\mathbf{y}$ is expanded in terms of its basis, where each basis vector is weighted by a temporal coefficient
\begin{equation}
	\mathbf{y} \approx \mathbf{U} \hat{\mathbf{y}}(t) \, ,
	\label{eq:galerkin expansion}
\end{equation}
where $\mathbf{\hat{y}(t)} \in \mathbb{R}^{r}$ is a vector of the temporal coefficients.
Second, the system is projected onto a POD subspace by multiplying the system on the left by $\mathbf{\Phi}^T$ and making use of the orthogonality of the POD basis (i.e, $\mathbf{U}^T\mathbf{U} = \mathbf{I}$), we end up with 
\begin{equation}
	\frac{d\mathbf{\hat{y}}}{dt} = \mathbf{A}_r\mathbf{\hat{y}} + \mathbf{\Phi}^T\mathbf{f}(\mathbf{U}\mathbf{\hat{y}}) \qquad \mathbf{\hat{y}}(0) = \mathbf{U}^T\mathbf{y}_0 \, ,
	\label{eq:ROM}
\end{equation}
where $\mathbf{A}_r \in \mathbb{R}^{r\times r} = \mathbf{\Phi}^T\mathbf{A}\mathbf{\Phi}.$
Thus, we are left with  $r$ ODE's instead of $n$ ODE's, which makes the resulting ROM more computationally efficien, and, once it is solved for $\mathbf{\hat{y}}$, the full order solution can be computed by Eq.~\ref{eq:galerkin expansion}.

%--------------------------------------------------------------------------------------------------%

\subsection{Discrete Empirical Interpolation Method}

\label{sec:DEIM}
Recently, the DEIM \cite{chaturantabut2010nonlinear} was proposed to approximate the non-linearity in model order reduction.
%The method is a discrete variant of the Empirical Interpolation Method (EIM) \cite{barrault2004empirical} that was originally proposed to approximate
%non-affine coefficient functions using a reduced-basis expansion to facilitate the online/offline decomposition in finite element methods.
The DEIM is proposed to facilitate the online/offline decomposition of ROMs.
The method relies on two main ingredients: (1) a reduced space that approximates the nonlinear function, and (2) a set of spatial interpolation indices.
To illustrate the method, let $\tau$ denote time or a model parameter.
The nonlinear function $\mathbf{f}(\textbf{y}(\tau))$ is approximated by projecting it onto a subspace that approximates the function space
\begin{equation}
	\mathbf{f}(\textbf{y}(\tau)) = \mathbf{\Psi}\boldsymbol{\theta}(\tau)\, ,
\end{equation}
where $\mathbf{\Psi} \in \mathbb{R}^{n\times k}$ is a low-dimensional subspace with $k \ll n$, and $\boldsymbol{\theta}(\tau) \in \mathbb{R}^k$ is a vector of the corresponding coefficients.
POD is used to find the subspace $\mathbf{\Psi}$ using snapshots of the function (i.e, $[\textbf{y}(\tau_1), ....., \textbf{y}(\tau_m)]$).
Note that these snapshots are computed at the same points used for the state variable snapshots required for the reduced model; hence this step will not introduce any additional cost other than computing the SVD.
In order to compute $\theta(\tau)$, a few spatial indices are selected iteratively using a greedy search following Algorithm \ref{alg:DEIM}.
The basic idea is to select $k$ rows of the basis matrix $\mathbf{\Psi}$.
The algorithm takes as an input the basis vectors of the subspace $\mathbf{\Psi}$.
The first index is selected where the entry of the first input basis has the largest magnitude.
Each other index $\mathcal{I}_i$ is selected where the approximation of the corresponding input basis vector $\mathbf{\Psi_i}$ in the subspace constructed by interpolating the basis $[\mathbf{\Psi}_1, ....\mathbf{\Psi}_{i-1}]$ at the previously selected indices is worst.
Once the indices are selected, the matrix $\mathbf{\Psi}_k \in \mathbb{R}^{k\times k}$ is formed such that
\begin{equation}
	\mathbf{\Psi}_k = \mathbf{P}^T\mathbf{\Psi}\, ,
\end{equation}
where $\mathbf{P} = [e_{z1}, e_{z2}, ....,e_{zk}] \in \mathbb{R}^{n\times k}$ and $e_{zi} = [0,.., 0, \underbrace{1}_{{zi}}, 0..., 0]^\text{T} \in \mathbb{R}^{n}$
is the  ${zi}$th column of the identity matrix.
The coefficients can be computed by solving
\begin{equation}
	\mathbf{P}^T\mathbf{f(\tau)}=\mathbf{\Psi}_k \boldsymbol{\theta}(\tau)\, ,
	\label{eq:DEIM coefficients}
\end{equation}
and the final function approximation is given by
\begin{equation}
	\hat{\mathbf{f}}(\tau)\approx\mathbf{\Psi}( \mathbf{\Psi}_k)^{-1}\mathbf{P}^T\mathbf{f} (\tau)\, .
\end{equation}
Note that the matrix $\mathbf{P}$ does not need to be constructed explicitly, since the term $\mathbf{P}^T\mathbf{f}$ extracts only the elements of $\mathbf{f}$ that correspond to the interpolation indices.
Thus, at each time step, only $k$ elements $\mathbf{f}$ need to be evaluated.
The approximation error can be bounded as 
\begin{equation}
	\|\mathbf{f} - \hat{\mathbf{f}} \| \le \|(\mathbf{P}\mathbf{\Psi})^{-1} \| \epsilon \, ,
\end{equation}
where
\begin{equation*}
	\epsilon \approx \sigma_{k+1} \, .
\end{equation*}
Note that $\sigma_{k+1}$ is the $(k+1)th$ singular value coming from the SVD of the function snapshots matrix.
Full derivation of the error bound is given in Ref.~\cite{chaturantabut2010nonlinear}.

As explained in the previous section, the projection can be performed by computing the term $\mathbf{U}^T \mathbf{\Psi}$ once in an offline stage, while in the online stage, the vector $\boldsymbol{\theta}$ is computed by solving the linear system of dimension $k$ in Eq.~\ref{eq:DEIM coefficients}.
The key point is that the cost of the online computation depends on the reduced dimension $k$, not the original, high dimension $n$.
The matrix version of the DEIM (MDEIM) is suggested in \cite{carlberg2012efficient, benner2015survey} to approximate parameter/time-dependent operators.
Consider a parameterized matrix $\mathbf{A}(\boldsymbol{\mu}) \in \mathbb{R}^{n\times n}$, the goal is to find a decomposition similar to Eq.~\ref{Eq:matrix decomposition} so that an offline-online ROM splitting is attainable.
The DEIM procedures can be exploited to find such decomposition. First, the matrix can be cast in a vector form by stacking all of its columns in one vector $\mathbf{a}(\mu) \in \mathbb{R}^{n^2}$.
This vector can be treated in the same exact way as the function $\mathbf{f}(t)$ in the previous section, i.e, $\mathbf{a}(\mu) = \mathbf{\Psi}\boldsymbol{\theta}(\mu)$.
Snapshots of the serialized matrix are collected at different parameter values and the reduced subspace $\mathbf{\Psi}$ and the interpolation indices are computed following algorithm \ref{alg:DEIM}.
Again, the corresponding coefficients are obtained by solving Eq.~\ref{eq:DEIM coefficients}. 
Next, each vector in $\mathbf{\Psi}$ can be mapped back to an $n \times n$ matrix, i.e, $\textbf{A}_i$.
Hence, the matrix is approximated as $\mathbf{\hat{A}}(t)=\sum_{i=0}^k \theta_i(t) \mathbf{A}_i$ and the projection term is computed as in Eq.~\ref{Eq:deim matrix projection}.
As pointed out in Ref.\cite{negri2015efficient}, the Weyl\textendash Mirsky theorem \cite{weyl1912asymptotische} ensures that the singular values of the approximated matrix approach that of the original as $k$ increases 
\begin{equation}
	|\sigma_i - \hat{\sigma_i}  |\ \le \|\mathbf{A} - \hat{\mathbf{A}}\|_\text{F}
\end{equation}
where $\sigma_i$ and $\hat{\sigma_i}$ denote the $i$th singular value of $\mathbf{A}$ and $\hat{\mathbf{A}}$ respectively.
Moreover, some properties of the original matrix such as positive definiteness can be preserved \cite{carlberg2015preserving}, if necessary, by imposing some constrains to the problem in Eq.~\ref{eq:DEIM coefficients}.

%In the practical implementations, the method efficiency can be improved, as suggested in Ref. \cite{negri2015efficient}.
%First, the matrix $\mathbf{A}$ is, generally, largely sparse, which leads to a snapshots matrix that is sparse, too.
%Consequently, it is more convenient to store only the nonzero elements ($nz$) in $\mathbf{a}$, 
%which should be appreciably more efficient in terms of memory and computational expense.
%In addition, to obtain the expansion coefficients ($\boldsymbol{\theta}$), the term $\mathbf{P}^T\mathbf{a(\tau)}$ needs to be computed at each new parameter value.
%In fact, this term only extracts the elements of $\mathbf{a(\tau)}$ that corresponds to the selected  interpolation indices.
In the offline stage, a mapping can be done between these interpolation indices and the corresponding row-column pairs in $\mathbf{A}$.
When the matrix is formed in the online stage, only these entries are evaluated by restricting a \textit{for loop } over these row-column pairs, leading to more reduction in the computational cost.

\noindent\begin{minipage}{\textwidth}
	\renewcommand\footnoterule{}
	\begin{algorithm}[H]
		\caption{DEIM}
		\label{alg:DEIM}
		\begin{algorithmic}[1]
			\REQUIRE Snapshots matrix: $\mathbf{X} = [\mathbf{f}_1, \mathbf{f}_2,  ...., \mathbf{f}_m] \in \mathbb{R}^{n\times m}$
			\ENSURE A reduced space $\mathbf{\Psi}$, set of indices $\mathcal{I}$
			\STATE $[\mathbf{\Psi}_1, \mathbf{\Psi}_2, ..., \mathbf{\Psi}_k ]$ = POD(${\mathbf{X}}, \epsilon$)\footnote{POD refers to Algorithm \ref{alg:pod}}
			\STATE $\mathcal{I}_1 = \argmaxH(\abs{\mathbf{\Psi}_1})$, $\mathcal{I}$ = $\{\mathcal{I}_1\}$
			\STATE $\mathbf{\Psi} = [\mathbf{\Psi}_1]$, $\mathbf{P} =[e_{I_1}]$
			\FOR {i=2 \TO i=k }
			\STATE Solve $(\mathbf{P}^T\mathbf{\Psi})\mathbf{c} = \mathbf{P}^T\mathbf{\Psi}_i$;
			\STATE $\mathcal{I}_i = \argmaxH{\abs{\mathbf{\Psi}_i - \mathbf{\Psi}\mathbf{c}}}$
			\STATE $\mathcal{I} = \mathcal{I} \cup \mathcal{I}_i$, $\mathbf{P} = \mathbf{P} \cup e_{I_i}$, ${\Psi = \Psi \cup \Psi_i}$
			\ENDFOR
		\end{algorithmic}
	\end{algorithm}
\end{minipage}

\subsection{Greedy POD}

A greedy sampling in the parameter space \cite{nguyen2009reduced} can be combined with POD to construct a global subspace that can be used in the framework of parametric ROMs.
The primary goal is to ensure that the parameter domain of interest (i.e, $\mathbb P$) is efficiently sampled and includes enough information such that it can be used or wide range of parameters including those of higher dimensions.
To construct this subspace, a training parameters dataset, $P_{train} \subset \mathbb P$, is prepared. 
The greedy subspace is initialized with the first $r_1$ POD basis computed from a randomly selected sample of the training data.
Next, this subspace is used to compute the ROM prediction error of the state variable for the rest of the training samples.
The sample at which the error is worst is selected, and the subspace is enriched by adding the first $r_1$ POD basis computed from the selected sample snapshots.
Selecting the samples with the worst errors ensures efficient sampling of the parameter space.
Intuitively, samples that are close to the previously chosen parameter points will have a smaller ROM error.
On the other hand, one expects that the ROM will fail to give an accurate prediction at samples that are far from the included points in the current subspace.
Thus, we keep adding new basis vectors until a maximum size of the subspace is reached or a user-specified error tolerance at a reference parameter point $\boldsymbol{\mu}_{\text{ref}}$ is met.
Note that each time a new basis vector is added, the subspace is orthonormalized.
Regarding the error criteria based on which a new sample is selected, there are two approaches. 
Ideally, the true error can be used leading to ``strong POD-greedy'' \cite{benner2017model}.
Alternatively, an error indicator can be used to alleviate the computational burden associated with evaluating the full-order-model. 
Here, we used the strong-POD greedy approach. 
The algorithm steps are shown in Algorithm \ref{alg:greedy pod}. 
The term $\Delta(\textbf{U},\boldsymbol{\mu}_\text{ref})$  is the error measure using the current reduced subspace at $\boldsymbol{\mu}_\text{ref}$, and $r$ is the maximum size of the greedy basis.

@@

%-----------------------------------------------------------------------------------------------------%
\section{Application}

\subsection{Time Dependent Diffusion Equation}
\label{sec:fom}
The time dependent diffusion equation was used as the mathematical model describing the neutronic transient in a nuclear reactor, i.e FOM and for which a an approximate ROM is sought.
It is represented as
\begin{equation}
	\begin{split}
	\frac{1}{v_g} \frac{\partial\phi_g}{\partial t} &= 
	\nabla. D_g\nabla\phi_g - \Sigma_{tg}\phi_g + \sum_{g^\prime=1}^{G}{\Sigma_{sg^\prime g}}\phi_{g^\prime} \\
	&+ \chi_g(1-\beta)\sum_{g^\prime = 1}^{G}\nu\Sigma_{fg^\prime}\phi_{g^\prime}  
	+ \chi_g\sum_{i=1}^{P} \lambda_i C_i \, ,
	\end{split}
	\label{time_diff}
\end{equation}
where 
\begin{equation*}
	\frac{\partial C_i}{\partial t} = -\lambda_i C_i + \beta_i \sum_{g^\prime=1}^{G} \nu\Sigma_{fg^\prime} \phi_{g^\prime} \, ,
\end{equation*}
and $\phi_g$, $D_g$, $\Sigma_{ag}$, $\nu\Sigma_{fg}$ and $v_g$ are the neutron flux, diffusion coefficient, the absorption cross section, the production cross section and the average neutron velocity in group g, respectively. $\Sigma_{sg^\prime g}$ is the scattering cross section from group $g^\prime$ to group $g$ 
and $\chi_g$ is the probability that a fission neutron will be born in group $g$.
$C_i$ is the $i$th precursor concentration, $\beta_i$ is the fraction of delayed neutrons in group $i$, $\lambda_i$ is the $i$th group decay constant.
The system can be represented compactly in a matrix form as 
\begin{equation}
\frac{\partial{y}(r, t)}{\partial{t}} = \mathcal{A}{y}(r, t) \, ,
\end{equation}
where the stacked unknowns are
\begin{equation*}
	{y} = 
	\begin{pmatrix}
	\phi_1(r, t) \\
	\vdots \\
	\phi_G(r, t)  \\
	{C}_1(r, t) \\
	\vdots \\
	{C}_I(r, t) \\
	\end{pmatrix} \, .
\end{equation*}

%\begin{equation*}
%\mathcal{A} = \left(\begin{array}{@{}c|c@{}}
%\begin{matrix}
%v_1(\nabla \cdot D_1\nabla +\Sigma_{11}) & v_1\Sigma_{12} & \hdots  &v_1\Sigma_{1G} \\
%v_2\Sigma_{21}        &    v_2(\nabla\cdot D_2\nabla +\Sigma_{22})            & \hdots  &v_2\Sigma_{2G} \\
%\vdots   &   \vdots   &  \ddots  &\vdots  \\
%v_G\Sigma_{G1}        &  v_G\Sigma_{G2} & \hdots &  v_G(\nabla \cdot D_G\nabla +\Sigma_{GG})     \\
%\end{matrix}
%&\begin{matrix}
%v_1 f_{11} & \hdots &   v_1 f_{1I} \\
%\vdots   & \ddots & \vdots \\
%v_G f_{G1} & \hdots &   v_G f_{GI}
%\end{matrix}
%\\
%\hline
%\begin{matrix}
%p_{11} &  &  & p_{12} & & & & \hdots  &   & p_{1G} \\
%\vdots &  &  & & \ddots & & & \vdots&   & \vdots \\
%p_{I1} &  &  & p_{I2} &  &   & & \hdots&  & p_{IG} 
%\end{matrix}
%&
%\begin{matrix}
%-\lambda_1 & \hdots & 0 \\
%\vdots & \ddots  &  \vdots \\
%0  &  \hdots  & -\lambda_I
%\end{matrix}
%\end{array}\right)
%\end{equation*}
%where 
%\begin{subequations}
%	\begin{equation*}
%	\Sigma{gg^\prime} = (1 -\beta)\chi_g\nu\Sigma{fg^\prime} + \Sigma_{s_{gg^\prime}} \, ,
%	\end{equation*}
%	\begin{equation*}
%	f_{ig^\prime} = \chi_g\lambda_i \, ,
%	\end{equation*}
%\end{subequations}
%\vspace{-2 pt}
%and
%\vspace{-2 pt}
%\begin{equation*}
%p_{ig} = \beta_i \nu\Sigma_{fg} \, .
%\end{equation*}
Upon spatial discretization, the system of ODEs can be compactly represented in a block matrix form as  
\begin{equation}
	\frac{d\mathbf{y}(t)}{dt} = \mathbf{A}\mathbf{y}(t) \, ,
	\label{eq:transient fom}
\end{equation}
where $\mathbf{y}\in\mathbb{R}^{N}$ is stacked vector of the flux and the precursors concentration, $\mathbf{A}\in\mathbb{R}^{N\times N}$ and $N = (I + G)\times n$ with $n$ is the number of spatial unknowns arising from discretization.
The discretized block matrix $\mathbf{A}$ is written as 
\begin{equation}
	\mathbf{A} =  \left(\begin{array}{@{}c|c@{}}
	\mathbf{L} & \mathbf{D} \\
	\hline
	\mathbf{F} & \mathbf{P}
	\end{array}\right) \, .
\end{equation} 
where $\mathbf{L} \in \mathbb{R}^{(n\times G)\times (n\times G)}$ represents the diffusion operator, $\mathbf{D} \in \mathbb{R}^{(n\times G)\times (n\times I)}$ represents the delayed neutrons production operator , $\mathbf{P} \in \mathbb{R}^{(n\times I)\times ((n\times G)}$ represents the precursors loss operator and $\mathbf{F}\in \mathbb{R}^{(n\times I)\times ((n\times I)}$ represents the precursors production operator.
%Thus, the system in Eq.~\ref{eq:transient fom} is the full-order model and it can be solved using a suitable time-scheme.
%However, it is expensive to solve such systems for fine spatial and time grids, since a linear system of a size proportional to the spatial discretization has to be solved for each time step.
%Hence, a reduced-order model to approximate it is justified.

%--------------------------------------------------------------------------------------%
\subsection{Benchmark Desription}

The LRA benchmark employs two energy groups and two delayed precursor groups and represents a 2-D quarter-core, BWR model subject to a control-rod ejection with adiabatic heating \cite{anl}.
A schematic layout of the problem is shown in Fig.~\ref{fig:lra core}, where region numbers indicate different materials and region R is the ejected rod.
%To produce time-dependent snapshots, the deterministic transport code Detran\footnote{https://github.com/robertsj/libdetran} was used to solve the time-dependent diffusion equations.
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=0.6]

	\begin{scope}<->;
